Galois Characterization of Function Fields of \\ Transcendence Degree $1$ over Number Fields

### 2017-06-26 Master Thesis/synopsis.tex
Recall Kummer theory facts

Let $K$ be a (not necessarily finitely generated) field over $\mathbb{Q}$ and of dimension $2$.

\begin{lemma}\phantomsection\label{lemma-defectless-valuations}
\begin{enumerate}[(i)]
\item $\operatorname{Spv}_\text{d}^{(1)}(K) = \operatorname{Spv}_\text{rd}^{(1)}(K)$
\item $\operatorname{Spv}^{(2)}(K) = \operatorname{Spv}_\text{d}^{(2)}(K)\subset |\operatorname{Spv}(K)|$
\end{enumerate}
\end{lemma}

\begin{proof}
\begin{enumerate}[(i)]
\item Let $v$ be a defectless valuation of rational rank $1$. By Lemma \ref{rank-dimension-inequality}, $\operatorname{rank}(v) \leq \operatorname{rrank}(v) = 1$. Since $\Gamma_v$ is not trivial, $\operatorname{rank}(v) = 1=\operatorname{rrank}(v)$.
\item Let $v$ be a valuation of rational rank $2$. Due to Lemma \ref{rank-dimension-inequality}, the dimension inequality is an equality and we have $\dim k(v) = 0$. \qedhere
\end{enumerate}
\end{proof}

>>>>

Let $K$ be a (not necessarily finitely generated) field over $\mathbb{Q}$ and of dimension $2$. First, let us recall some facts from Galois cohomology:

\begin{proposition}[\textit{Kummer Theory}]\label{prop:kummer-theory}
Let $F$ be a field and $n\in\mathbb{N}$ invertible in $F$. Let $\mu_n\subset F^\text{s}$ be the group of all $n$-th roots of unity. Consider the following short exact sequence:
\[ 1\longrightarrow \mu_n\longrightarrow (F^\text{s})^\times \stackrel{n}{\longrightarrow} (F^\text{s})^\times \longrightarrow 1 \]
This gives rise to the following long exact cohomology sequence:
\[\begin{tikzcd}
\cdots\longrightarrow F^\times \ar[r, "n"]& F^\times \ar[r]& \mathrm{H}^1(F, \mu_n)\ar[r] & \mathrm{H}^1(F, (F^\text{s})^\times)\ar[dlll]\\
\mathrm{H}^1(F, (F^\text{s})^\times)\ar[r] & \mathrm{H}^2(F, \mu_n)\ar[r] & \operatorname{Br}(F)\ar[r, "n"] & \operatorname{Br}(F) \longrightarrow \cdots
\end{tikzcd}\]
Since $\mathrm{H}^1(F, (F^\text{s})^\times)=0$ by Hilbert's Theorem 90, we get isomorphisms:
\begin{align*}
\mathrm{H}^1(F,\mu_n) \cong F^\times/n, \qquad \mathrm{H}^2(F, \mu_n) \cong {}_n\operatorname{Br}(F)
\end{align*}
\end{proposition}
